Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach

### 2013-06-25 inst/doc/nonparametric-bayes.tex
writing

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{A <- myers_posteriors}
\NormalTok{A$index <- A$index + A$chain * }\KeywordTok{max}\NormalTok{(A$index) }\CommentTok{# Combine samples across chains by renumbering index }
\NormalTok{myers_pardist <- }\KeywordTok{acast}\NormalTok{(A, index ~ variable)}
\NormalTok{bayes_coef <- }\KeywordTok{apply}\NormalTok{(myers_pardist,}\DecValTok{2}\NormalTok{, posterior.mode) }\CommentTok{# much better estimates}
\NormalTok{myers_bayes_pars <- }\KeywordTok{unname}\NormalTok{(}\KeywordTok{c}\NormalTok{(bayes_coef[}\DecValTok{2}\NormalTok{], bayes_coef[}\DecValTok{3}\NormalTok{], bayes_coef[}\DecValTok{1}\NormalTok{]))}
\NormalTok{myers_means <- }\KeywordTok{sapply}\NormalTok{(x_grid, Myer_harvest, }\DecValTok{0}\NormalTok{, myers_bayes_pars)}
\NormalTok{myers_f <- function(x,h,p) }\KeywordTok{Myer_harvest}\NormalTok{(x, h, p[}\KeywordTok{c}\NormalTok{(}\StringTok{"r0"}\NormalTok{, }\StringTok{"theta"}\NormalTok{, }\StringTok{"K"}\NormalTok{)])}
\KeywordTok{head}\NormalTok{(myers_pardist)}
\end{Highlighting}
\end{Shaded}

@@

\begin{verbatim}
[1] 35.4576  0.5129 23.9914
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{allen_deviance <- }\KeywordTok{posterior.mode}\NormalTok{(pardist[,}\StringTok{'deviance'}\NormalTok{])}
\NormalTok{ricker_deviance <- }\KeywordTok{posterior.mode}\NormalTok{(ricker_pardist[,}\StringTok{'deviance'}\NormalTok{])}
\NormalTok{myers_deviance <- }\KeywordTok{posterior.mode}\NormalTok{(myers_pardist[,}\StringTok{'deviance'}\NormalTok{])}
\NormalTok{true_deviance <- }\DecValTok{2}\NormalTok{*}\KeywordTok{estf}\NormalTok{(}\KeywordTok{c}\NormalTok{(p, sigma_g))}
\NormalTok{mle_deviance <- }\DecValTok{2}\NormalTok{*}\KeywordTok{estf}\NormalTok{(}\KeywordTok{c}\NormalTok{(est$p, est$sigma_g))}

\KeywordTok{c}\NormalTok{(}\DataTypeTok{allen =} \NormalTok{allen_deviance, }\DataTypeTok{ricker=}\NormalTok{ricker_deviance, }\DataTypeTok{myers=}\NormalTok{myers_deviance, }\DataTypeTok{true=}\NormalTok{true_deviance, }\DataTypeTok{mle=}\NormalTok{mle_deviance)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
  allen  ricker   myers    true     mle 
  32.93   35.92   35.46  -61.08 -889.99 
\end{verbatim}

@@

Decision-theoretic methods often rely on simple parametric models of
ecological dynamics to compare the value of a potential sequence of
actions. Unfortunately, such simple models rarely capture the complexity
or uncertainty found in most real ecosystems. Non-parametric Bayesian
methods offer a promising statistical approach for predictive modeling
of ecological dynamics in regions of state space where the data is
adequate, while at the same time offering more flexible patterns with
greater uncertainty outside the observed data. This contrasts from
simple parametric models which provide relatively constant level of
uncertainty in regions with and without adequate data. The consequence
of such misplaced confidence outside the data can lead to highly
undesirable results that may be avoided with the more flexible
non-parametric Bayesian approach.

@@

Most management recommendations from the ecological literature are based
on (or at least motivated by) parametric models. Though in principle
these models reflect mechanistic underpinnings of biological
interactions involved, in practice real mechanisms are often unknown and
the true dynamics too complex to be captured by simple models. While
simple mechanistic models can nevertheless provide important insights
into possible dynamics -- for instance, demonstrating that a vaccine
does not have to be 100\% effective to eliminate the transmission of a
virus -- such approaches are not well-suited for use in forecasting
outcomes of potential management options. Non-parametric approaches
offer a more flexible alternative that can both more accurately reflect
the data available while also representing greater uncertainty in areas
(of state-space) where data is lacking. Ecological research and
management strategy should pay closer attention to the opportunities and
challenges nonparametric modeling can offer.

\subsubsection{Quantitative vs Qualitative Decisions}

In this paper, we consider those ecological management problems in which
a mathematical (or computational) model is used to quantitatively inform
decision-making by allowing a manager to compare to the expected
consequences of potential management actions (or policies). We
distinguish this from the solely qualitative use of a model, in which
models are used to represent and compare hypotheses of different
mechanisms that would lead to qualitatively different actions.

Though simple mechanistically motivated models may be best for the
latter case (@Cuddington2013; Geritz and Kisdi 2012), such models can be
not only inaccurate but misleading in quantitative decision making. This
arises for two reasons:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  We do not know what the correct parameter values are for the models.
\item
  We do not know what the correct models are for ecological systems.
\end{enumerate}

The complexity of ecological interactions and a lack of data contribute
greatly to both of the problems. This concern is particularly acute in
case of ecological tipping points (Barnosky et al. 2012; Scheffer et al.
2001), which arise from feedbacks common in ecological systems and can
lead to a sudden catastrophic transition to an undesirable state. In
general we do not know where such tipping points are unless we have
already observed the transition, in which case it is frequently too late
to respond. Though there may exist early warning signals for certain
kinds of these transitions that are driven by slow changes (Scheffer et
al. 2009), we do not know when, where, or how to apply them to the
decision making context more generally (Boettiger and Hastings 2013).

In addition to facing the potentially grave consequences of such
transitions, if they do exist in a system of interest, this generally
means that we lack data in the region near and beyond a possible tipping
point. As a consequence, our models are least accurate just where we
need them to be most accurate.\\Unfortunately, parametric modeling
approaches are not well-suited to this scenario, in which good data is
avialble over only part of the relevant state-space. Necessary
assumptions about model structure in any parametric approach can lead
such models to perform very poorly, even when accounting for parameter
uncertainty.

Non-parametric models can more realistically represent uncertainties
outside of observed data while also better capturing the dynamics in the
region observed. Advances in the theory and computational
implementations of nonparametric methods make them ripe for such
applications. We use the classic problem of optimal harvest of a marine
fishery to illustrate how the nonparametric approach of Gaussian
processes can be applied. We will compare Bayesian implementations of
both nonparametric and parametric models, which best allow us to capture
the uncertainty of model estimation in either case and permits a more
natural comparison between approaches.

@@

\subsubsection{Uncertainties: the classical approach}

The ecological dynamics of most systems of management interest are
typically both complex and unknown. Despite this, quantitative policy
optimization in these systems has almost exclusively been based on
simple, mechanistically motivated parametric models. There are several
reasons for this. First, limited available data almost always precludes
any accurate estimate of more complicated models. Second, the
computational demands of determining an optimal strategy in a sequential
decision-theory problem suffer from the ``curse of dimensionality'' that
makes it difficult to consider more complicated models, states, or
actions. Unfortunately, these challenges also make it particularly
difficult to reflect the true uncertainty either in the parameters of
the given models (parametric uncertainty) or in the structure of the
models themselves (structural uncertainty). Further unknowns such as
measurement uncertainty, parameter uncertainty, unobserved states,
knowledge of boundary conditions, etc. further compound the issue.
Though a hierarchical Bayesian approach provides a natural way to
address these from a statistical standpoint, formulating reasonable
parametric descriptions of each form of uncertainty is a challenging
task in itself, let alone the computational difficulties of solving such
a system. Cressie et al. (2009) provides a good account of the successes
and challenges of the approach. Applying these approaches in the
management context of sequential decision making, in which forecasts
must be obtained over a range of possible actions and updated regularly
as new information arrives makes such an approach less feasible still.

An active literature and growing computational power over the past
several decades have only marginally improved this situation. Parametric
or structural uncertainty can be introduced only by increasing the state
space to reflect a distribution of possible parameters or the degree of
belief in each of a set of possible models Frequently, the space of
possible actions must then be reduced or the algorithms adjusted by
approximations to keep computations feasible.

Traditional approaches to optimal control (Pontryagin's principle,
stochastic dynamic programming) rely on knowledge of the state equation,
usually described by a simple parametric model. Here we illustrate how a
stochastic dynamic programming algorithm can alternatively be driven by
the predictions from a Gaussian process.

@@

\begin{align}
X_{t+1} &= Z_t f(S_t) \\
S_t &= X_t - h_t \\
f(S_t) &= S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
\end{align}

@@

In our example, we focus on the problem in which a manager must set the
harvest level for a marine fishery each year to maximize the net present
value of the resource, given an estimated stock size from the year
before. Rich data and global concerns have made marine fisheries the
crucible for much of the founding work (Gordon 1954; Reed 1979; May et
al. 1979; Ludwig and Walters 1982) in managing ecosystems under
uncertainty. Global trends (Worm et al. 2006) and controversy (Hilborn
2007; Worm et al. 2009) have made understanding these challenges all the
more pressing.

@@

A crux of this approach is correctly specifying the functional form of
$f$, along with its parameters. The standard approach uses one of a
handful of common parametric models representing the stock-recruitment
relationship, usually after estimating the model parameters from any
available existing data. Uncertainty in the parameter estimates can be
estimated and integrated over to determine the optimal policy under
under uncertainty (Mangel 1985; Schapaugh and Tyre 2013). Uncertainty in
the model structure itself can only be addressed in this approach by
hypothesizing alternative model structures, and then performing some
model choice or model averaging (B. K. Williams 2001; Athanassoglou and
Xepapadeas 2012). For simplicity of the comparison we consider only the
case of a structurally correct and structurally incorrect model, and
estimate Bayesian posterior distributions of the parameters. We will
also compare to a simpler maximum likelihood estimate of the parameters
that ignores the issue of parameter uncertainty.

We compare these parametric approaches to our alternative that uses a
Gaussian Process (GP) in place of a given stock recruitment curve. We
briefly revisit a description of the Gaussian Process approach that has
appeared in recent literature before describing how we take the
important step of adapting the GP to the decision theory problem that
can still be solved using SDP.

@@

The use of Gaussian process (GP) regression (or ``kreging'' in the
geospatial literature) to formulate a predictive model is relatively new
in the context of modeling dynamical systems (Kocijan et al. 2005) and
introduced in the ecological modeling and fisheries management by Munch
et al. (2005). An accessible and thorough introduction to the
formulation and use of GPs can be found in Rasmussen and Williams
(2006).

@@

Barnosky, Anthony D., Elizabeth a Hadly, Jordi Bascompte, Eric L.
Berlow, James H. Brown, Mikael Fortelius, Wayne M. Getz, et al. 2012.
``Approaching a state shift in Earth's biosphere.'' \emph{Nature} 486
(7401) (jun): 52--58. doi:10.1038/nature11018.
\url{http://www.nature.com/doifinder/10.1038/nature11018}.

@@

Boettiger, Carl, and Alan Hastings. 2013. ``Tipping points: From
patterns to predictions.'' \emph{Nature} 493 (7431) (jan): 157--158.
doi:10.1038/493157a.
\url{http://www.nature.com/doifinder/10.1038/493157a}.

@@

Cressie, Noel, Catherine a Calder, James S. Clark, Jay M. Ver Hoef, and
Christopher K. Wikle. 2009. ``Accounting for uncertainty in ecological
analysis: the strengths and limitations of hierarchical statistical
modeling.'' \emph{Ecological Applications} 19 (3) (apr): 553--70.
\url{http://www.ncbi.nlm.nih.gov/pubmed/19425416}.

@@

Geritz, Stefan A. H., and Eva Kisdi. 2012. ``Mathematical ecology: why
mechanistic models?.'' \emph{Journal of mathematical biology} 65 (6-7)
(dec): 1411--5. doi:10.1007/s00285-011-0496-3.
\url{http://www.ncbi.nlm.nih.gov/pubmed/22159789}.

@@

Scheffer, Marten, Jordi Bascompte, William A. Brock, Victor Brovkin,
Stephen R. Carpenter, Vasilis Dakos, Hermann Held, Egbert H. van Nes,
Max Rietkerk, and George Sugihara. 2009. ``Early-warning signals for
critical transitions.'' \emph{Nature} 461 (7260): 53--9.
\url{http://www.ncbi.nlm.nih.gov/pubmed/19727193}.

>>>>

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{A <- myers_posteriors}
\NormalTok{A$index <- A$index + A$chain * }\KeywordTok{max}\NormalTok{(A$index) }\CommentTok{# Combine samples across chains by renumbering index }
\NormalTok{myers_pardist <- }\KeywordTok{acast}\NormalTok{(A, index ~ variable)}
\NormalTok{bayes_coef <- }\KeywordTok{apply}\NormalTok{(myers_pardist,}\DecValTok{2}\NormalTok{, posterior.mode) }\CommentTok{# much better estimates}
\NormalTok{myers_bayes_pars <- }\KeywordTok{unname}\NormalTok{(}\KeywordTok{c}\NormalTok{(bayes_coef[}\StringTok{"r0"}\NormalTok{], bayes_coef[}\StringTok{"theta"}\NormalTok{], bayes_coef[}\StringTok{"K"}\NormalTok{]))}
\NormalTok{myers_means <- }\KeywordTok{sapply}\NormalTok{(x_grid, Myer_harvest, }\DecValTok{0}\NormalTok{, myers_bayes_pars)}
\NormalTok{myers_f <- function(x,h,p) }\KeywordTok{Myer_harvest}\NormalTok{(x, h, p[}\KeywordTok{c}\NormalTok{(}\StringTok{"r0"}\NormalTok{, }\StringTok{"theta"}\NormalTok{, }\StringTok{"K"}\NormalTok{)])}
\KeywordTok{head}\NormalTok{(myers_pardist)}
\end{Highlighting}
\end{Shaded}

@@

\begin{verbatim}
[1]  0.5129  1.0894 23.9914
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{allen_deviance <- -}\DecValTok{2}\NormalTok{*}\KeywordTok{posterior.mode}\NormalTok{(pardist[,}\StringTok{'deviance'}\NormalTok{])}
\NormalTok{ricker_deviance <- -}\DecValTok{2}\NormalTok{*}\KeywordTok{posterior.mode}\NormalTok{(ricker_pardist[,}\StringTok{'deviance'}\NormalTok{])}
\NormalTok{myers_deviance <- -}\DecValTok{2}\NormalTok{*}\KeywordTok{posterior.mode}\NormalTok{(myers_pardist[,}\StringTok{'deviance'}\NormalTok{])}
\NormalTok{true_deviance <- }\DecValTok{2}\NormalTok{*}\KeywordTok{estf}\NormalTok{(}\KeywordTok{c}\NormalTok{(p, sigma_g))}
\NormalTok{mle_deviance <- }\DecValTok{2}\NormalTok{*}\KeywordTok{estf}\NormalTok{(}\KeywordTok{c}\NormalTok{(est$p, est$sigma_g))}
\NormalTok{xtable::}\KeywordTok{xtable}\NormalTok{(}\KeywordTok{as.table}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DataTypeTok{Allen =} \NormalTok{allen_deviance, }\DataTypeTok{Ricker=}\NormalTok{ricker_deviance, }\DataTypeTok{Myers=}\NormalTok{myers_deviance, }\DataTypeTok{True=}\NormalTok{true_deviance, }\DataTypeTok{MLE=}\NormalTok{mle_deviance)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Thu Jun 20 16:29:00 2013
\begin{table}[ht]
\centering
\begin{tabular}{rr}
  \hline
 & x \\ 
  \hline
Allen & -65.85 \\ 
  Ricker & -71.85 \\ 
  Myers & -70.92 \\ 
  True & -61.08 \\ 
  MLE & -889.99 \\ 
   \hline
\end{tabular}
\end{table}
\end{verbatim}

@@

Decision-theoretic methods often rely on simple parametric models of
ecological dynamics to compare the value of a potential sequence of
actions. Unfortunately, such simple models rarely capture the complexity
or uncertainty found in most real ecosystems.

We demonstrate how nonparametric Bayesian models can provide robust,
nearly-optimal solutions to decision making under uncertainty when
\emph{we don't know the correct model} to use.

While methods that account for \emph{parametric} uncertainty can be very
successful with the right model, structural uncertainty of not knowing
what model best approximates the dynamics poses considerably greater
difficulty.

Non-parametric Bayesian methods offer a promising statistical approach
for predictive modeling of ecological dynamics in regions of state space
where the data is adequate, while at the same time offering more
flexible patterns with greater uncertainty outside the observed data.
This contrasts from simple parametric models which provide relatively
constant level of uncertainty in regions with and without adequate data.
The consequence of such misplaced confidence outside the data can lead
to highly undesirable results that may be avoided with the more flexible
non-parametric Bayesian approach.

@@

Decision making under uncertainty is a ubiquitous challenge of natural
resource management and conservation.

Here we illustrate how a stochastic dynamic programming algorithm can be
driven by the predictions from a Gaussian process model, sidestepping
the need for an accurate model-based description of the sustem dynamics.

Decision-theoretic (or optimal control) tools require a model that can
assign probabilities of future states (e.g.~stock size of a fishery)
given the current state and a proposed action (e.g.~fishing harvest or
effort). The decision maker then seeks to determining the course of
actions (also referred to as the policy) that maximizes the expected
value of some objective function (such as net present value derived from
the resource over time. (The approach can be adapted to permit
alternatives to maximizing expectation, such as minimizing the maximum
cost or damage that might be incurred. See @Polasky2011). Management
frequently faces a sequential decision-making problem -- after selecting
an action, the decision-maker may receive new information about the
current state and must again choose an appropriate action --
i.e.~setting the harvest limits each year based on stock assessments the
year prior.

Traditional approaches to optimal control (Pontryagin's principle,
stochastic dynamic programming) rely on knowledge of the state equation.
Such simple parametric models, frequently justified by mechanistic
underpinnings are used to provide probabilities of future states.
Fisheries literature, which enjoys some of the longest history in the
use of the optimal control framework, provides a clear example. Models
such as the Ricker or Beverton Holt curves are ubiquitous {[}too strong?
references?{]} Parametric models can reflect variability introduced by a
stochastic environment or demographic process my incorporating random
shocks from a given distribution.

As the parameter values for these models must be estimated from limited
data,\\there will always be some uncertainty associated with these
values. This uncertainty further compounds the intrinsic variability
introduced by demographic or environmental noise. The degree of
uncertainty in the parameter values can be inferred from the data and
reflected in the estimates of the transition probabilities.

Unfortunately, estimates of parameter uncertainty are only as good as
the parametric models themselves. Too often, we do not understand the
system dynamics well enough to know if a model provides a good
approximation over the relevant range of states and timescales (criteria
that we loosely refer to as defining the ``right'' or ``true'' model.)
So called structural or model uncertainty is a more difficult problem
than parametric uncertainty. Typical solutions involve either model
choice, or model averaging. Either approach remains limited by the
degree to which any of the proposed models are sufficiently close to the
right model.

The nature of decision-making problems provides a convenient way to
compare models. Rather than compare models in terms of best fit to data
or fret over the appropriate penalty for model complexity, model
performance is defined in the concrete terms of the decision-maker's
objective function, which we will take as given. (Much argument can be
made over the `correct' objective function, e.g.~how to account for the
social value of fish left in the sea vs.~the commercial value of fish
harvested; see @Halpern2013 for further discussion of this issue.
Alternatively, we can always compare model performance across multiple
potential objective functions.) The decision-maker does not necessarily
need a model that provides the best mechanistic understanding or the
best long-term outcome, but rather the one that best estimates the
probabilities of being in different states as a result of the possible
actions.

While simple mechanistic models can nevertheless provide important
insights into long-term outcomes, such approaches are not well-suited
for use in forecasting outcomes of potential management options.
Non-parametric approaches offer a more flexible alternative that can
both more accurately reflect the data available while also representing
greater uncertainty in areas (of state-space) where data is lacking.

We demonstrate how a Gaussian Process model of stock recruitment can
lead to nearly optimal management through stochastic dynamic
programming, comperable to knowing the correct structural equation for
the underlying simulation. Meanwhile, parametric models that do not
match the underlying dynamics can perform very poorly, even though they
fit the data as well as the true model.

Ecological research and management strategy should pay closer attention
to the opportunities and challenges nonparametric modeling can offer.

Non-parametric models can better represent uncertainties outside of
observed data while also better capturing the dynamics in the region
observed. Advances in the theory and computational implementations of
nonparametric methods make them ripe for such applications. We use the
classic problem of optimal harvest of a marine fishery to illustrate how
the nonparametric approach of Gaussian processes can be applied. We will
compare Bayesian implementations of both nonparametric and parametric
models, which best allow us to capture the uncertainty of model
estimation in either case and permits a more natural comparison between
approaches.

@@

@@

\[X_{t+1} = Z_t f(S_t)  \] \[S_t = X_t - h_t \]
\[f(S_t) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)} \]

@@

In our example, we focus on the problem in which a manager must set the
harvest level for a marine fishery each year to maximize the net present
value of the resource, given an estimated stock size from the year
before. The economic value and ecological concern have made marine
fisheries the crucible for much of the founding work (Gordon 1954; Reed
1979; May et al. 1979; Ludwig and Walters 1982) in managing ecosystems
under uncertainty. Global trends (Worm et al. 2006) and controversy
(Hilborn 2007; Worm et al. 2009) have made understanding these
challenges all the more pressing.

@@

A crux of this approach is correctly specifying the functional form of
$f$, along with its parameters. The standard approach uses one of a
handful of common parametric models representing the stock-recruitment
relationship, usually after estimating the model parameters from any
available existing data. Uncertainty in the parameter estimates can be
estimated and integrated over to determine the optimal policy under
under uncertainty (Mangel 1985; Schapaugh and Tyre 2013). Uncertainty in
the model structure itself can only be addressed in this approach by
hypothesizing alternative model structures, and then performing some
model choice or model averaging (B. K. Williams 2001; Athanassoglou and
Xepapadeas 2012).

\subsubsection{Parametric Models}

We consider three candidate parametric models for the stock-recruitment
function, which we refer to by the first authors of the publications in
which they were first proposed.

We generate the data with a four-parameter model that contains a tipping
point (an Allee effect, see {[}@Allen, @Courchamp{]}) below which the
stock decreases to zero,

\[ X_{t+1} = Z_t S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(\frac{S_t - \theta}{K}\right)} \]

\[ S_t = X_t - h_t \]

The parameter $C$ reflects the location of the tipping point, $K$ the
carrying capacity of the stock, and $r$ the base recruitment rate. $S_t$
represents the stock size after a harvest $h_t$ has been implemented.
$Z_t$ represents a log-normal random variable of log-mean zero and
log-standard deviation parameter $\sigma$.

We consider two alternative candidate models: the Ricker {[}@Ricker{]}
stock-recruitment curve,

\[X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \]

and an alternative four-parameter model adapted from @Myers,

\[ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \]

which contains a tipping point for $\theta > 2$ and becomes a
Beverton-Holt model at $\theta = 1$.

\subsubsection{Bayesian Inference of Parametric models}

Given the sample data, we infer posterior distributions for each of the
three models listed above using a Markov Chain Monte Carlo Gibbs Sampler
(jags, see appendix for implementation details and code) given uniform
priors. We run six chains for $10^6$ steps each and then assess
convergence by Gelman-Rubin criterion and inspection of the traces, see
appendix.

@@

The use of Gaussian process (GP) regression (or ``kriging'' in the
geospatial literature) to formulate a predictive model is relatively new
in the context of modeling dynamical systems (Kocijan et al. 2005) and
introduced in the ecological modeling and fisheries management by Munch
et al. (2005). An accessible and thorough introduction to the
formulation and use of GPs can be found in Rasmussen and Williams
(2006).

@@

@@

@@

@@

@@
